\documentclass{article}

% Language setting
\usepackage[english]{babel}

% Set page size and margins
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{enumitem}
\usepackage{float}

\title{Web Application Threat Model}
\author{Aditya Pangavhane}

\begin{document}
\maketitle

\begin{abstract}
Cybersecurity is a discipline that demands both technical expertise and strategic foresight. The process of securing digital assets begins long before vulnerabilities are discovered or exploited; it starts with a deep understanding of the system, its environment, and the evolving threat landscape. Threat modeling is the cornerstone of this proactive approach, enabling organizations to anticipate how adversaries might target their systems and to design defenses that are both robust and adaptable.

This report provides a comprehensive exploration of threat modeling methodologies, including STRIDE, PASTA, Trike, VAST, OCTAVE, and OWASP. Each framework is examined in detail, with practical examples and case studies that illustrate their application in real-world scenarios. The document also integrates hands-on labs using Linux security tools, demonstrating how theoretical concepts translate into actionable security practices.

Throughout the chapters, readers will find in-depth discussions on the evolution of threat modeling, technical definitions, and the integration of security controls into the software development lifecycle. The report emphasizes the importance of collaboration between technical and non-technical stakeholders, highlighting how effective communication and continuous improvement are essential for maintaining a resilient security posture.

By connecting technical threats to business risks—such as reputational damage, regulatory penalties, and operational downtime—this report aims to bridge the gap between cybersecurity professionals and decision-makers. The actionable recommendations and best practices provided herein are designed to empower organizations to build secure systems, foster a culture of security, and stay ahead of emerging threats in an increasingly complex digital world.
\end{abstract}


% =====================
% Expanded Chapter Structure
% =====================

\section{Introduction}
\input{chapters/01_introduction}

\section{Background and Evolution of Threat Modeling}
\input{chapters/02_background}

\section{STRIDE: Microsoft’s Threat Modeling Framework}
\input{chapters/03_stride}

\section{PASTA: Process for Attack Simulation and Threat Analysis}
\input{chapters/04_pasta}

\section{Other Frameworks: Trike, VAST, OCTAVE, and OWASP}
\input{chapters/05_other_frameworks}

\section{Threat Modeling Methodology: Step-by-Step}
\input{chapters/06_methodology}

\section{Case Study: Threat Model of a Vulnerable Web Application}
\input{chapters/07_case_study}

\section{Security Controls, Mitigations, and Best Practices}
\input{chapters/08_controls}

\section{Risk Assessment, Reporting, and Continuous Improvement}
\input{chapters/09_risk_reporting}

\section{Lab: Practical Threat Modeling with Linux Commands}
\input{chapters/10_lab}

\section{Conclusion and Future Directions}
\input{chapters/11_conclusion}

\section{References}
\input{chapters/12_references}

\end{document}
